---
title: "Getting Started with Was2CoDE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Was2CoDE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Getting Started with Was2CoDE

The Was2CoDE package provides a tool for analyzing gene expression data on individual-level using IDEAS mehtod and decomposed components of Wasserstein-2 distance. This vignette demonstrates how to use the package functions to analyze synthetic gene expression data.

## Installation

First, ensure that you have installed the Was2CoDE package from GitHub:

```{r install, eval=FALSE}
# Don't run these during build/check
install.packages("devtools")
install.packages("data.table")
install.packages("doRNG")

# Install Was2CoDE from GitHub (example only)
devtools::install_github("TatiZhang/Was2CoDE")
```


## Load the package

```{r}
# Load the packages
library(Was2CoDE)
library(data.table)
library(doRNG)
```

## Generate Synthetic Data
```{r}
set.seed(0)

# Number of genes and cells
n_genes <- 10
n_cells <- 20

# Generate synthetic gene expression data
count_input <- matrix(rnorm(n_genes * n_cells, mean = 5, sd = 2), nrow = n_genes, ncol = n_cells)
rownames(count_input) <- paste0("gene", 1:n_genes)
colnames(count_input) <- paste0("cell", 1:n_cells)

# Generate metadata for cells
meta_cell <- data.frame(
  cell_id = colnames(count_input),
  individual = rep(paste0("donor", 1:5), each = n_cells / 5),
  some_variable = rnorm(n_cells)  # Add a placeholder variable for var_per_cell
)

# Generate metadata for individuals
meta_ind <- data.frame(
  individual = unique(meta_cell$individual),
  var2test = rnorm(5)
)

var_per_cell <- "some_variable"
var2test <- "var2test"

```

## Analyze Gene Expression Data
```{r}
# Compute distance metrics
# result <- Was2CoDE::was2code_dist(count_input, meta_cell, meta_ind, var_per_cell, var2test)
# 
# # Print the result
# print(result)

```


