---
title: "Getting Started with IdeasCustom"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with IdeasCustom}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Getting Started with IdeasCustom

The IdeasCustom package provides a tool for analyzing gene expression data on individual-level using IDEAS mehtod and decomposed components of Wasserstein-2 distance. This vignette demonstrates how to use the package functions to analyze synthetic gene expression data.

## Installation

First, ensure that you have installed the IdeasCustom package from GitHub:

```{r}
# Install devtools if you haven't already
install.packages("devtools")

# Install IdeasCustom from GitHub
devtools::install_github("TatiZhang/IdeasCustom")
```

## Load the package

```{r}
# Load the package
library(IdeasCustom)
```

## Generate Synthetic Data
```{r}
set.seed(123)

# Number of genes and cells
n_genes <- 10
n_cells <- 20

# Generate synthetic gene expression data
count_input <- matrix(rnorm(n_genes * n_cells, mean = 5, sd = 2), nrow = n_genes, ncol = n_cells)
rownames(count_input) <- paste0("gene", 1:n_genes)
colnames(count_input) <- paste0("cell", 1:n_cells)

# Generate metadata for cells
meta_cell <- data.frame(
  cell_id = colnames(count_input),
  individual = rep(paste0("donor", 1:5), each = n_cells / 5)
)

# Generate metadata for individuals
meta_ind <- data.frame(
  individual = unique(meta_cell$individual),
  var2test = rnorm(5)
)

var_per_cell <- "some_variable"
var2test <- "var2test"

```



